Peer-to-Peer Permissionless Consensus via Reputation

### 2023-08-11 freechains/freechains.tex
...

In an ideal Internet forum, all messages or posts
(i)   reach all users,
(ii)  are delivered in consistent order, and
(iii) are respectful and on topic.
In centralized systems, items (i) and (ii) are trivially achieved assuming
availability and delivery order, while item (iii) requires that users trust the
service to moderate content.
In decentralized systems, however, none of these demands are easily
accomplished.
A common approach in gossiping protocols is to proactively disseminate posts
among peers until they reach all users~\cite{p2p.survey,p2p.byz}.
However, gossiping does not guarantee consensus since posts can be received in
conflicting orders~\cite{p2p.intention}.
%,p2p.dvcs}.
%As an example, antagonistic messages such as \emph{"X is final"} vs
%\emph{"Y is final"} might be sent concurrently, preventing the network to
%determine as a group its intention as \emph{X} or \emph{Y}.
%
Consensus is key to distinguish malicious behavior \emph{at the protocol
level}, and therefore, satisfy item (iii).
Without consensus, each client must rely on local anti-abuse policies, such as
SPAM filters and blacklists, but which only apply a posteriori, when the
abusive messages have already been flooded in the network by potential
Sybils~\cite{p2p.sybil}.
%
%fundamental in collaborative applications, such as open discussion
%forums, chat rooms, and trading services (e.g., delivery \& auction).
%Otherwise, they become impractical due to excess and abuse (e.g., SPAM and
%illicit content).
%At the very core, consensus is

@@

Bitcoin~\cite{p2p.bitcoin} is the first permissionless protocol to resist
Sybil attacks, relying on a scarce resource---the \emph{proof-of-work}---to
establish consensus.
%The protocol maintains a single dispendious timeline consisting of linked
%blocks with transactions, which represent the consensus.
%Alternative timelines need to spend more resources to substitute the consensus.
The protocol is Sybil resistant because it is expensive to write to its
unique timeline (either via proof-of-work or transaction fees).
%
However, Bitcoin and cryptocurrencies in general are not suitable for social
content sharing because
    (i)   they enforce a unique timeline to preserve value and immunity to
          attacks;
    (ii)  they lean towards concentration of power due to scaling effects;
    (iii) they impose an external economic cost to use the protocol; and
    (iv)  they rely exclusively on objective rules to reach consensus.
%
These issues threaten our original decentralization goals, and more
importantly, they renounce subjectivity, which is inherent to content
moderation.
%
For instance, a unique timeline implies that all Internet content should be
subject to the same consensus rules, while objective rules alone cannot
distinguish abuse amid content.
%
Another limitation of cryptocurrencies is that it is not possible to revoke
content in the middle of a blockchain, which is inadmissible considering
illicit content (e.g., hate speech)~\cite{btc.content}.
%For instance, Bitcoin transactions permanently store arbitrary human-authored
%content, including harmful texts and images.
%~\cite{btc.data}

In this work, we adapt the idea of scarce resources to reach consensus and
eradicate Sybils, but in the context of social content sharing.
We indicate 4 main contributions:
%
\begin{enumerate}
\item Recognizing published contents as scarce resources, since they require
human work.
Work is manifested as new posts, which if approved by others, reward authors
with reputation tokens, which are further used to evaluate other posts with
likes and dislikes.
With such \emph{proof-of-authoring} mechanism, token generation is expensive,
while verification is cheap and shared by multiple users.
%
\item Using the reputation system to determine consensus.
Due to decentralization and network partitions, posts in a timeline form a
causal graph with only partial order, which we promote to a total order based
on the reputation of authors.
The consensus order is fundamental to detect conflicting operations, such as
likes with insufficient reputation (akin to Bitcoin's double spending).
%
\item Allowing users to create diversified forums of interest (instead of a
singleton blockchain), each counting as an independent timeline with its own
subjective authoring etiquette.
%
\item Supporting content removal without compromising the integrity of the
decentralized blockchain.
Users have the power to revoke posts with dislikes, and peers are forced to
remove payloads, only forwarding associated blockchain metadata.
\end{enumerate}
%
We integrated the proposed consensus algorithm into \FC~\cite{fcs.sbseg20}, a
practical peer-to-peer (P2P) content dissemination protocol that provides
strong eventual consistency~\cite{p2p.sec}.
%p2p.crdts,
To show the feasibility of the consensus mechanism, we simulated months of
activity of a chat channel and years of a newsgroup forum, both extracted from
publicly available Internet archives.

>>>>

In an ideal Internet forum, all messages or posts
(i)   reach all users,
(ii)  are delivered in consistent order, and
(iii) are respectful and on topic.
In centralized systems, items (i) and (ii) are trivially achieved assuming
availability and delivery order, while item (iii) requires users to trust the
service policies.
In decentralized systems, however, none of these demands are easily
accomplished.
A common approach in gossiping protocols is to proactively disseminate posts
among peers until they reach all users~\cite{p2p.survey,p2p.byz}.
However, gossiping does not guarantee consensus since posts can be received in
conflicting orders~\cite{p2p.intention}.
%,p2p.dvcs}.
%As an example, antagonistic messages such as \emph{"X is final"} vs
%\emph{"Y is final"} might be sent concurrently, preventing the network to
%determine as a group its intention as \emph{X} or \emph{Y}.
%
Consensus is key to distinguish malicious behavior \emph{at the protocol
level}, and therefore, satisfy item (iii).
Without consensus, each client must rely on local anti-abuse policies, such as
SPAM filters, but which only apply a posteriori, when the abusive messages have
already been flooded in the network by potential Sybils~\cite{p2p.sybil}.
%
%fundamental in collaborative applications, such as open discussion
%forums, chat rooms, and trading services (e.g., delivery \& auction).
%Otherwise, they become impractical due to excess and abuse (e.g., SPAM and
%illicit content).
%At the very core, consensus is

@@

Bitcoin~\cite{p2p.bitcoin} is the first permissionless protocol to resist
Sybils, relying on a scarce resource---the \emph{proof-of-work}---to establish
consensus.
%The protocol maintains a single dispendious timeline consisting of linked
%blocks with transactions, which represent the consensus.
%Alternative timelines need to spend more resources to substitute the consensus.
The protocol is Sybil resistant because it is expensive to write to its
unique timeline (either via proof-of-work or transaction fees).
%
However, Bitcoin and cryptocurrencies in general are not suitable for social
content sharing because
    (i)   they enforce a unique timeline to preserve value and immunity to
          attacks;
    (ii)  they lean towards concentration of power due to scaling effects;
    (iii) they impose an external economic cost to use the protocol; and
    (iv)  they rely exclusively on objective rules to reach consensus.
%
These issues threaten our decentralization goals, and more importantly, they
renounce subjectivity, which is inherent to content moderation.
%
For instance, a unique timeline implies that all Internet content is subject
to the same consensus rules, while objective rules alone cannot distinguish
abuse amid content.
%
Another limitation of cryptocurrencies is that it is not possible to revoke
content in the middle of a blockchain, which is inadmissible considering
illicit content (e.g., hate speech)~\cite{btc.content}.
%For instance, Bitcoin transactions permanently store arbitrary human-authored
%content, including harmful texts and images.
%~\cite{btc.data}

In this work, we adapt the idea of scarce resources to reach consensus and
eradicate Sybils, but in the context of social content sharing.
We indicate 4 main contributions:
%
\begin{enumerate}
\item Recognizing published contents as scarce resources, since they require
human work.
Work is manifested as new posts, which if approved by others, reward authors
with reputation tokens, which are further used to evaluate other posts with
likes and dislikes.
With such \emph{proof-of-authoring} mechanism, token generation is expensive,
while verification is cheap and shared by multiple users.
%
\item Using the reputation system to determine consensus.
Due to decentralization and network partitions, posts in a timeline form a
causal graph with only partial order, which we promote to a total order based
on the reputation of authors.
The consensus order is fundamental to detect conflicting operations, such as
likes with insufficient reputation (akin to Bitcoin's double spending).
%
\item Allowing users to create diversified forums of interest (not a
singleton blockchain), each as an independent timeline with its own
subjective authoring etiquette.
%
\item Supporting content removal without compromising the integrity of the
decentralized blockchain.
Users have the power to revoke posts with dislikes, and peers are forced to
remove payloads, only forwarding associated blockchain metadata.
\end{enumerate}
%
We integrated the proposed consensus algorithm into \FC~\cite{fcs.sbseg20}, a
practical peer-to-peer (P2P) content dissemination protocol that provides
strong eventual consistency~\cite{p2p.sec}.
%p2p.crdts,
To show the feasibility of the consensus mechanism, we simulated months of
activity of a chat channel and years of a newsgroup forum, both extracted from
publicly available Internet archives.
